---
title: "exercise-8-EEB177"
author: "Mehdi Nojoumi"
date: "March 1, 2017"
output: md_document
---
```{r}
rm(list=ls())
.First <- function() {
  options(width=70)
  options(useFancyQuotes=FALSE)
  require(ascii)
  require(xtable)
  require(ggplot2)
  require(scales)
  require(reshape2)
  require(plyr)
  theme_set(theme_bw())
  png <<- function(res=96, width=500, height=300, ...) grDevices::png(res=res, width=width, height=height, ...)
}

```

```{r}
housing <- read.csv("dataSets/landdata-states.csv")
head(housing[1:5])


hist(housing$Home.Value)
```


```{r}
library(ggplot2)
ggplot(housing, aes(x = Home.Value)) +
  geom_histogram()
```

```{r}
plot(Home.Value ~ Date,
     data=subset(housing, State == "MA"))
points(Home.Value ~ Date, col="red",
       data=subset(housing, State == "TX"))
legend(1975, 400000,
       c("MA", "TX"), title="State",
       col=c("black", "red"),
       pch=c(1, 1))
```

```{r}
ggplot(subset(housing, State %in% c("MA", "TX")),
       aes(x=Date,
           y=Home.Value,
           color=State))+
  geom_point()
```

```{r}
help.search("geom_", package = "ggplot2")
```

```{r}
hp2001Q1 <- subset(housing, Date == 2001.25) 
ggplot(hp2001Q1,
       aes(y = Structure.Cost, x = Land.Value)) +
  geom_point()
```

```{r}
ggplot(hp2001Q1,
       aes(y = Structure.Cost, x = log(Land.Value))) +
  geom_point()
```

```{r}
hp2001Q1$pred.SC <- predict(lm(Structure.Cost ~ log(Land.Value), data = hp2001Q1))

p1 <- ggplot(hp2001Q1, aes(x = log(Land.Value), y = Structure.Cost))

p1 + geom_point(aes(color = Home.Value)) +
  geom_line(aes(y = pred.SC))

```

```{r}
p1 +
  geom_point(aes(color = Home.Value)) +
  geom_smooth()
```

```{r}
p1 + 
  geom_text(aes(label=State), size = 3)

```

```{r}
#install.packages("ggrepel") 
library("ggrepel")
p1 + 
  geom_point() + 
  geom_text_repel(aes(label=State), size = 3)
```


```{r}
p1 +
  geom_point(aes(size = 2),# incorrect! 2 is not a variable
             color="red") # this is fine -- all points red
```

```{r}
p1 +
  geom_point(aes(color=Home.Value, shape = region))
```

# Exercise I:
***
1-Create a scatter plot with CPI on the x axis and HDI on the y axis.
```{r}
library("ggrepel")
library(ggplot2)
dat <- read.csv("dataSets/EconomistData.csv")
head(dat)

ggplot(dat, aes(x = CPI, y = HDI)) + geom_point()

```

2-Color the points blue.
```{r}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point(color="Blue")
```

3-Map the color of the the points to Region.
```{r}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point(aes(color=Region))
```

4-Make the points bigger by setting size to 2
```{r}
ggplot(dat, aes(x = CPI, y = HDI, size = 2)) + geom_point(aes(color=Region))
```

5-Map the size of the points to HDI.Rank
```{r}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point( aes(color=Region, size= HDI.Rank))
```



```{r}
args(geom_histogram)
args(stat_bin)
```

```{r}
p2 <- ggplot(housing, aes(x = Home.Value))
p2 + geom_histogram()
```

```{r}
p2 + geom_histogram(stat = "bin", binwidth=4000)
```

```{r}
housing.sum <- aggregate(housing["Home.Value"], housing["State"], FUN=mean)
rbind(head(housing.sum), tail(housing.sum))
```

```{r}
ggplot(housing.sum, aes(x=State, y=Home.Value)) + 
  geom_bar(stat="identity")
```



# Exercise II:
***
1-Re-create a scatter plot with CPI on the x axis and HDI on the y axis (as you did in the previous exercise).
```{r}
library("ggrepel")
library(ggplot2)
dat <- read.csv("dataSets/EconomistData.csv")

ggplot(dat, aes(x = CPI, y = HDI)) + geom_point()
```


2-Overlay a smoothing line on top of the scatter plot using geom_smooth.
```{r}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point() +
  geom_smooth()
```

3-Overlay a smoothing line on top of the scatter plot using geom_smooth, but use a linear model for the predictions. Hint: see ?stat_smooth.
```{r}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point() +
  geom_smooth(method = "lm", formula = y ~ x)

```

4-Overlay a smoothing line on top of the scatter plot using geom_line. Hint: change the statistical transformation.
```{r}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point() +
  stat_smooth(stat="loess")

```

5-BONUS: Overlay a smoothing line on top of the scatter plot using the default loess method, but make it less smooth. Hint: see ?loess.
```{r}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point() +
  geom_smooth(formula = y ~ x, span=0.3)

```
<br><br>

```{r}

p3 <- ggplot(housing,
             aes(x = State,
                 y = Home.Price.Index)) + 
        theme(legend.position="top",
              axis.text=element_text(size = 6))
(p4 <- p3 + geom_point(aes(color = Date),
                       alpha = 0.5,
                       size = 1.5,
                       position = position_jitter(width = 0.25, height = 0)))
```

```{r}
p4 + scale_x_discrete(name="State Abbreviation") +
  scale_color_continuous(name="",
                         breaks = c(1976, 1994, 2013),
                         labels = c("'76", "'94", "'13"))
```

```{r}
p4 +
  scale_x_discrete(name="State Abbreviation") +
  scale_color_continuous(name="",
                         breaks = c(1976, 1994, 2013),
                         labels = c("'76", "'94", "'13"),
                         low = "blue", high = "red")

```

```{r}
p4 +
  scale_color_continuous(name="",
                         breaks = c(1976, 1994, 2013),
                         labels = c("'76", "'94", "'13"),
                         low = ("blue"), high = ("red"))

```

```{r}
p4 +
  scale_color_gradient2(name="",
                        breaks = c(1976, 1994, 2013),
                        labels = c("'76", "'94", "'13"),
                        low = ("blue"),
                        high = ("red"),
                        mid = "gray60",
                        midpoint = 1994)
?muted
```

# Exercise III:

1.Create a scatter plot with CPI on the x axis and HDI on the y axis. Color the points to indicate region.
```{r}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point(aes(color=Region))
```


2-Modify the x, y, and color scales so that they have more easily-understood names (e.g., spell out "Human development Index" instead of "HDI").
```{r}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point(aes(color=Region))+scale_x_continuous(name="Corruption Perception Index")+scale_y_continuous(name="Human development Index")
```

3-Modify the color scale to use specific values of your choosing. Hint: see ?scale_color_manual.
```{r}
cols <- c("Americas"= "purple","Asia Pacific"= "black","East EU Cemt Asia"= "white","EU W. Europe"= "blue","MENA"= "red","SSA"= "darkgreen")
ggplot(dat, aes(x = CPI, y = HDI)) +   geom_point(aes(color=Region))+scale_x_continuous(name="Corruption Perception Index")+scale_y_continuous(name="Human development Index")+scale_color_manual(values = cols)

``` 


```{r}
p5 <- ggplot(housing, aes(x = Date, y = Home.Value))
p5 + geom_line(aes(color = State))
```

```{r}
(p5 <- p5 + geom_line() +
   facet_wrap(~State, ncol = 10))
```

```{r}
p5 + theme_linedraw()
```

```{r}
p5 + theme_light()
```

```{r}
p5 + theme_minimal() +
  theme(text = element_text(color = "turquoise"))

#opts.png

```

```{r}
theme_new <- theme_bw() +
  theme(plot.background = element_rect(size = 1, color = "blue", fill = "black"),
        text=element_text(size = 12, family = "Serif", color = "ivory"),
        axis.text.y = element_text(colour = "purple"),
        axis.text.x = element_text(colour = "red"),
        panel.background = element_rect(fill = "pink"),
        strip.background = element_rect(fill = ("orange")))

p5 + theme_new
```

```{r}
housing.byyear <- aggregate(cbind(Home.Value, Land.Value) ~ Date, data = housing, mean)
ggplot(housing.byyear,
       aes(x=Date)) +
  geom_line(aes(y=Home.Value), color="red") +
  geom_line(aes(y=Land.Value), color="blue")
```

```{r}
library(tidyr)
home.land.byyear <- gather(housing.byyear,
                           value = "value",
                           key = "type",
                           Home.Value, Land.Value)
ggplot(home.land.byyear,
       aes(x=Date,
           y=value,
           color=type)) +
  geom_line()

```

